A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1} and low-resolution spectroscopy

### 2017-06-29 paper/comoving-followup.tex
introduction

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}

@@

\begin{abstract}
% Context
Widely-separated, coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and measuring the
mass distribution of the Milky Way.
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
The kinematic properties of bound wide binaries and their unbound complements
serve as potentiometers for the smooth tidal field and perturbation spectrum in
the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
However, without radial velocity measurements for the vast majority of the
pairs, the contamination rate at large separations is expected to be
considerable.
% Methods
Here, we present results from our own radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the XX comoving pairs observed, we find YY ...
% Conclusions
Gaia DR2 -> how many do we expect?
\end{abstract}

@@

Outline:
- Stars form in clusters, chemically homogeneous at some level
- Clusters generally disrupt slowly through a process that preserves initial
  clustering in conserved quantities but destroys phase coherence
- Most significant process is probably tidal heating
- Dissolved clusters are therefore useful dynamical tracers of mass distribution
- In Galactic halo, globular clusters
- In disk, hope is to use chemical signatures to connect stars, then can use
  dynamics to reconstruct birth clusters, infer mass distribution / evolution
- Clusters have advantage of many stars per cluster, but fewer clusters
- Wide binary stars are little potentiometers themselves - only N=2, but are far
  more numerous as a population
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?
- Also useful for stellar models: should have same age, chemistry, but need to
  test this
    - Chemical homogeneity tested with few open clusters - wide binaries another
      test
    - Need sample of confirmed widely-separated binaries, follow-up to get
      detailed abundances
    - Need way to verify that widely separated stars are physically associated,
      via kinematics
- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\

@@

Most stars form in clusters with small dispersions in position, velocity, and
(likely) small initial spreads in chemical abundances.
Once a given cluster is visibly dispersing or fully dispersed, the former member
stars retain chemical and kinematic signatures of their birth cluster

@@

, though
the kinematic mapping is non-trivial because of phase-mixing.

become
useful

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{tgas}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
For a full description of the selection method, see \citealt{Oh:2017}.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above; the likelihood ratio is computed as
$L_1/L_2$.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical
velocity components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by
computing the same likelihood ratio for random pairings of stars, we find
13,058 comoving star pairs with physical separations between $10^3~\au$ and
$10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.

>>>>

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}

@@

\begin{abstract}
% Context
Widely-separated, coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
The kinematic properties of bound wide binaries and their unbound complements
serve as potentiometers for the smooth tidal field and perturbation spectrum in
the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
However, without radial velocity measurements for the vast majority of the
pairs, the contamination rate at large separations is expected to be
considerable.
% Methods
Here, we present results from our own radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the XX comoving pairs observed, we find YY ...
% Conclusions
Gaia DR2 -> how many do we expect?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same age and initial chemistry
(e.g., cite).
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way; obvious dynamical
associations between stars from the same cluster are then erased as the
constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way.
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children so that they may be re-associated with precisely measured
abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on the intrinsic chemical
dispersions of stars formed from the same birth clusters, the precision of the
abundance measurements, and the ability to tell that the chemical abundances of
two stars are more similar than two randomly-paired Galactic disk stars.
Recent work suggests that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
It was also recently shown that $\approx 0.3\%$ of random pairings of Galactic
field red giant stars have abundance patterns that are indistinguishable at the
same level (``doppelgangers''), even with 20 chemical species
(\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied? And do the chemical properties of open clusters depend on the
star formation rate or efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses, and member stars that span a large range in
masses and evolutionary stages.
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated, coeval pairs or multiplets of stars.
Summary of wide binary literature

- Also useful for stellar models: should have same age, chemistry, but need to
  test this
    - Chemical homogeneity tested with few open clusters - wide binaries another
      test
    - Need sample of confirmed widely-separated binaries, follow-up to get
      detailed abundances
    - Need way to verify that widely separated stars are physically associated,
      via kinematics
- Tests of chemical enrichment due to planetary systems

In this work, we present a catalog of genuine comoving pairs.
We expect high-resolution follow-up to study chemistry...

% Caution that prior results are for red giants, MS stars may have planetary
% imprints that get erased as dredge-up and convective layer erases...

Outline:
- Stars form in clusters, chemically homogeneous at some level
- Clusters generally disrupt slowly through a process that preserves initial
  clustering in conserved quantities but destroys phase coherence
- Most significant process is probably tidal heating
- Dissolved clusters are therefore useful dynamical tracers of mass distribution
- In Galactic halo, globular clusters
- In disk, hope is to use chemical signatures to connect stars, then can use
  dynamics to reconstruct birth clusters, infer mass distribution / evolution
- Clusters have advantage of many stars per cluster, but fewer clusters
- Wide binary stars are little potentiometers themselves - only N=2, but are far
  more numerous as a population
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?

- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\

@@

@@

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{tgas}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above; the likelihood ratio is computed as
$L_1/L_2$.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical
velocity components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by
computing the same likelihood ratio for random pairings of stars, we find
13,058 comoving star pairs with physical separations between $10^3~\au$ and
$10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.
For a full description of the selection method, see \citealt{Oh:2017}.
